---
title: "Mini introduction to R"
output: html_document
date: "2023-05-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a mini introduction to how to use R. The goal is to learn the general aspects of how R works to use Seurat.

# Running basic commands
```{r}
1 + 1 # arithmetic operators
```

```{r}
9 > 10 # logical operator for comparisons
```

# Variables 
Variables can be assigned using '<-' or '='

## Scalar
Used to save a single value
```{r}
a <- 1 + 0.02 # numeric
a
```

## Vectors
Used to save multiple values of the same type in a 1-D array
```{r}
vectorNumbers <- c(1, 2, 4) # Create a numeric vector containing numbers
vectorNumbers
```

```{r}
vectorStrings <- c("neuron", "glia") # Create a vector containing characteres
vectorStrings
```

## Matrix
Used to save multiple values of the same type in a 2-D array
```{r}
MatrixNum <- matrix(data = c(1,1,1,2,2,2,3,3,3,4,4,4), nrow = 3, ncol = 4) # Numeric matrix filled by column (3x4)
MatrixNum
```

## Data frame
Used to save multiple values of the same size in a 2-D array
```{r}
DataFrame <- data.frame( numbers = c(1,2,3), region = c("cortex", "hippocampus", "spinal cord"),
              cells = c(1000, 30488, 2084))

DataFrame
```

## List
Use to save different types of values of different sizes in a multidimensional array
```{r}
ListRandom <- list(DataFrame, MatrixNum, vectorStrings, vectorNumbers, a)
ListRandom
```


# Functions
Functions are blocks of code grouped together to handle similar inputs and produce a specific output.

For example, we can calculate the mean of a numeric vector
```{r}
mean(c(90,4,5,.3))
```

or sort it
```{r}
sort(c(90,4,5,.3))
```

get dimensions
```{r}
DataFrame
dim(DataFrame)
rownames(DataFrame)
colnames(DataFrame) 
```

# Access variables
'[]' lets you access variables using indices
```{r}
vectorGenes <- c("Foxp2", "Spp1", "Sdk1", "Satb1", "Calb")
vectorGenes [3] # get element in the 3rd position
vectorGenes [1:3] # get elements from 1st to 3rd position
vectorGenes [c(2,5)] # get element in the 2nd and 5th positions
```

Accessing matrices
```{r}
MatrixNum <- matrix(data = c(1,1,1,2,2,2,3,3,3,4,4,4), nrow = 3, ncol = 4) 
MatrixNum [, 1] # accessing all rows in the 1st column
MatrixNum [2, ] # accessing all columns in the 2nd row
```

Accessing data frame
```{r}
DataFrame <- data.frame( numbers = c(1,2,3,4), region = c("cortex", "hippocampus", "spinal cord", "cortex"), cells = c(1000, 30488, 2084, 8888))

DataFrame$region #access elements in column "region"
DataFrame$region [2] # access 2nd element in the column "region" 
DataFrame$region == "cortex" # logical indices that meet a condition
DataFrame [DataFrame$region == "cortex", ] # access rows that a meet condition
```

# Importing Data
 To analyze data in Seurat you will need to read gene expression data into R. The easiest way is to use the *read.table* function.   
 
Here, we read a table which contains a header and each element is separated by a ','. 
```{r, eval=FALSE}
mydata <- read.table("c:/mydata.csv", header=TRUE, sep=",")
```

For variable formats, take a look at the *read.table* help. 
```{r}
help("read.table")
```

# libraries

There are some variables type and function that are specific for some Libraries. You can install them in the "packages" section. Once installed, libraries need to be re-load every time you reset R.

```{r}
library("ggplot2") #load library "ggplot2"
data(PlantGrowth) # load sample data
head (PlantGrowth) # take a look at the first element of the data set
ggplot(data = PlantGrowth, aes(x = group, y = weight)) + geom_boxplot()

```

